A systematic search for transiting planets in the \KT\ data

### 2015-03-17 ms.tex
moving figures and some responses

In this \paper\ we present a data-analysis technique for exoplanet search and
characterization that is insensitive to spacecraft-induced trends in the light
curves.
We assume that the dominant trends in the observed light curves in each star
are caused by the spacecraft and are, therefore, shared with other stars.
We reduce the dimensionality by running PCA on stellar light curves to obtain
the dominant modes.
The search for planets proceeds by modeling the data as a linear combination
of 150 of these basis vectors and a transit model.
Our method builds on the ideas behind previous data-driven de-trending
procedures such as the \kepler\ pipeline pre-search data conditioning
\citep[\pdc;][]{Stumpe:2012, Smith:2012}, but (because of our simultaneous
fitting approach) we can use a much more flexible systematics model without
worrying about over-fitting.

@@

The key insight in our transit search method that sets it apart from the
other standard procedures is that no de-trending is necessary.
Instead, we can fit for the noise (or trends) and exoplanet signals
simultaneously.
This is theoretically appealing because it will be more sensitive to low
signal-to-noise transits.
The main reason for this is that the signal is never precisely orthogonal to
the systematics and any de-trending will over-fit.
This will, in turn, decrease the amplitude of the signal and distort its
shape.
In order to reduce this effect, most procedures use a very rigid model for the
trends.
For \KT, this rigidity has been implemented by effectively asserting that
centroid measurements contain all of the information needed to describe the
trends \citep{Vanderburg:2014, Aigrain:2015, Crossfield:2015}.
In the \kepler\ pipeline, this is implemented by allowing only a small number
of PCA components to contribute to the fit in the \pdc\ procedure.
Instead, we will use a large number of ELCs---a very flexible model---and use
a simultaneous fitting and marginalization to avoid over-fitting.

Physically, the motivation for our model---and the \pdc\ model---is that every
star on the detector should be affected by the same set of systematic effects.
These are caused by things like pointing jitter, temperature variations, and
other sources of PSF modulation.
Each of these effects will be imprinted in the light curves of many stars with
varying amplitudes and signs as a result of the varying flat field and PSF.
Therefore, while it is hard to write down a physical generative model for the
systematics, building a data-driven model might be possible.
This intuition is also exploited by other methods that model the systematics
using only empirical centroids \citep{Vanderburg:2014, Armstrong:2014,
Aigrain:2015, Crossfield:2015}, but our more flexible model should capture a
wider range of effects, including those related to PSF and temperature.
For example, \Fig{corr} shows the application of our model---with 150
ELCs---to a light curve with no known transit signals and the precision is
excellent.
If, however, we were to apply this systematics model alone (without a
simultaneous fit of the exoplanet transit model) to a light curve with
transits, we would be at risk of over-fitting and decreasing the amplitude of
the signal.

In our pipeline we simultaneously fit for the transit signal and the trends
using a rigid model for the signal and a relatively flexible model for the
systematic noise.
Specifically, we model the light curve as being generated by linear
combination of 150 ELCs and a ``box'' transit model at a specific period,
phase, and duration.
The mathematical details are given in \app{math}, but in summary, since the
model is linear, we can analytically compute the likelihood
function---conditioned on a specific period, phase, and duration---for the
depth \emph{marginalizing out the systematics model}.
The signal-to-noise of this depth measurement can then be used as a quality of
fit metric or candidate selection scalar.
This computation is expensive but, as described in the following \sectionname
s, it is possible to scale the method to a \KT-size dataset.

@@

@@

To do first-pass vetting for blended EBs among our catalog of
planetary candidates, we test for significant centroid offsets using
the machinery that we have already established for modeling the
systematic trends in the data, inspired by the methods used to vet
\kepler\ candidates \citep{Bryson:2013}.  This is only an initial vetting step
and a more complete characterization of our catalog's reliability is
forthcoming (Montet, Foreman-Mackey, \etal\ in preparation).

>>>>

In this \paper\ we present a data-analysis technique for exoplanet search and
characterization that is insensitive to spacecraft-induced trends in the light
curves.
We assume that the dominant trends in the observed light curves in each star
are caused by the spacecraft and are, therefore, shared with other stars.
We reduce the dimensionality by running PCA on stellar light curves to obtain
the dominant modes.
The search for planets proceeds by modeling the data as a linear combination
of 150 of these basis vectors and a transit model.
Our method builds on the ideas behind previous data-driven de-trending
procedures such as the \kepler\ pipeline pre-search data conditioning
\citep[\pdc;][]{Stumpe:2012, Smith:2012}, but (because of our simultaneous
fitting approach) we can use a much more flexible systematics model
\response{while being less prone to over-fitting}.

@@

The key insight in our transit search method that sets it apart from most
standard procedures is that no de-trending is necessary.
Instead, we can fit for the noise (or trends) and exoplanet signals
simultaneously.
This is theoretically appealing because it should be more sensitive to low
signal-to-noise transits \response{and similar methods have been shown to be
effective for finding transits in ground-based surveys \citep{Berta:2012}}.
The main motivation for this model is that the signal is never precisely
orthogonal to the systematics and any de-trending will over-fit.
This will, in turn, decrease the amplitude of the signal and distort its
shape.
In order to reduce these effects, most de-trending procedures use a very rigid
model for the systematics.
For \KT, this rigidity has been implemented by effectively asserting that
centroid measurements contain all of the information needed to describe the
trends \citep{Vanderburg:2014, Aigrain:2015, Crossfield:2015}.
In the \kepler\ pipeline, this is implemented by allowing only a small number
of PCA components to contribute to the fit in the \pdc\ procedure.
Instead, we will use a large number of ELCs---a very flexible model---and use
a simultaneous fitting and marginalization to avoid over-fitting.

Physically, the motivation for our model---and the \pdc\ model---is that every
star on the detector should be affected by the same set of systematic effects.
These are caused by things like pointing jitter, temperature variations, and
other sources of PSF modulation.
Each of these effects will be imprinted in the light curves of many stars with
varying amplitudes and signs as a result of the varying flat field and PSF.
Therefore, while it is hard to write down a physical generative model for the
systematics, building a data-driven model might be possible.
This intuition is also exploited by other methods that model the systematics
using only empirical centroids \citep{Vanderburg:2014, Armstrong:2014,
Aigrain:2015, Crossfield:2015}, but our more flexible model should capture a
wider range of effects, including those related to PSF and temperature.
For example, \Fig{corr} shows the application of our model---with 150
ELCs---to a light curve with no known transit signals and the precision is
excellent.

If we were to apply this systematics model alone (without a simultaneous fit
of the exoplanet transit model) to a light curve with transits, we would be at
risk of over-fitting and decreasing the amplitude of the signal.
\response{\Fig{overfitting} demonstrates this effect on a synthetic transit
injected into the light curve of a typical bright star.
The middle two panels in this Figure show the light curve de-trended using 10
and 150 ELCs respectively.
When only 10 ELCs are used, the measured transit depth is relatively robust
but this model is clearly not sufficient for removing the majority of the
systematic trends.
The model with 150 ELCs does an excellent job of removing the systematics but
it also distorts the transit shape and decreases the measured transit depth,
hence reducing the signal strength in the \project{BLS} spectrum
\citep{Kovacs:2002}.}

In our pipeline we simultaneously fit for the transit signal and the trends
using a rigid model for the signal and a relatively flexible model for the
systematic noise.
Specifically, we model the light curve as being generated by linear
combination of 150 ELCs and a ``box'' transit model at a specific period,
phase, and duration.
The mathematical details are given in \app{math}, but in summary, since the
model is linear, we can analytically compute the likelihood
function---conditioned on a specific period, phase, and duration---for the
depth \emph{marginalizing out the systematics model}.
The signal-to-noise of this depth measurement can then be used as a quality of
fit metric or candidate selection scalar.
This computation is expensive but, as described in the following \sectionname
s, it is possible to scale the method to a \KT-size dataset.
\response{The bottom panel of \Fig{overfitting} shows the application of this
joint transit--systematics model to the synthetic transit discussed
previously.
When the joint model is used, the correct transit depth is measured---the
transit is not distorted---but the systematics are also well-described by the
model.}

@@

\begin{figure}[p]
\begin{center}
\includegraphics[width=0.5\textwidth]{figures/overfitting.pdf}
\end{center}
\caption{%
\response{%
A comparison between de-trending using different numbers of ELCs and a
simultaneous fit of the systematics and transit model.
\emph{(a)} The raw photometry for EPIC 201374602 with a synthetic transit
injected at 65 days. In all plots, the photometry is measured in
parts-per-thousand (ppt).
\emph{(b)} The black points show the photometry de-trended using a linear
combination of 10 ELCs and the green line shows the true transit model.
The maximum likelihood transit depth is computed following \project{BLS}
\citep{Kovacs:2002}.
While some of the systematics are removed by this model, there is still a lot
of residual noise.
\emph{(c)} The same plot as panel \emph{(b)} but using 150 ELCs to de-trend.
This model removes the majority of the systematics but also distorts the
transit and weakens the signal; it reduces the measured transit depth.
\emph{(d)} The final panel shows the results of simultaneously fitting for
the transit and the systematics using 150 ELCs.
The maximum likelihood depth is computed as described in \app{math}.
Like panel \emph{(c)}, this model removes most of the systematics but does
not distort the transit or reduce the measured transit depth.
}
\figlabel{overfitting}}
\end{figure}

@@

To do first-pass vetting for blended EBs among our catalog of
planetary candidates, we test for significant centroid offsets using
the machinery that we have already established for modeling the
systematic trends in the data, inspired by the methods used to vet
\kepler\ candidates \citep{Bryson:2013}.  This is only an initial vetting step
and a more complete characterization of our catalog's reliability is
forthcoming (Montet, \etal\ in preparation).
